---
title: "Explorative data analysis - let's play around a bit"
author: "Philipp Bayer"
date: "2020-09-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup}
library(tidyverse)
library(knitr)
```

# Introduction

Here i make some of the tables for the manuscript.

```{r}
files_to_make_tables_for <- list.files(path = './data', pattern = '*lst')
```

Let's print the counts of R-gene classes for each type, that will be table 1:

```{r results='asis',  warning=FALSE, message=FALSE}
all_results <- c()
class_dict <- list()

for(i in seq_along(files_to_make_tables_for)) {
  
  f <- files_to_make_tables_for[i]
  if (f == 'Lee.preRGA.candidates.by.Blast.lst' ) {next}
  if (f == 'Lee.RGA.candidates.lst') {next}
  print(f)
  fh <- read_tsv(paste('./data/', f, sep=''), col_names = c('Name','Class','Type'))
  if (ncol(fh) == 2) {
    # weird change in behaviour of tidyverse- used to be all NA by default
    # when there were only 2 columns
    fh$Type <- NA
  }
  fh <- fh %>% unite(United, Class, Type)
  # remove NAs in files with 2 columns
  fh$United <- gsub(pattern = '_NA', replacement = '', fh$United)
  
  fh <- fh %>% mutate(Name = gsub('.t1', '', Name))
  
  class_dict[[gsub('Lee.|.candidates.lst','',f)]] <- fh
  all_results <- c(all_results, table(fh$United))

}

kable(enframe(all_results, name='Class', value = 'Count'))
```



Let's calculate the per-class gene variability.

```{r}
pav_table <- read_tsv('./data/soybean_pan_pav.matrix_gene.txt.gz')
#t_pav_table <- as_tibble(cbind(nms = names(pav_table), t(pav_table)))
t_pav_table <- pav_table %>% pivot_longer(cols= -1) %>% pivot_wider(names_from = "Individual",values_from = "value") %>% rename('Individual'='name')
```

```{r}
names <- c()
presences <- c()

for (i in seq_along(t_pav_table)){
  if ( i == 1) next
  thisind <- colnames(t_pav_table)[i]
  pavs <- t_pav_table[[i]]
  presents <- sum(strtoi(pavs))
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
res_tibb <- new_tibble(list(names = names, presences = presences))
```

res_tibb now stores for each gene, in how many individuals it is present. We have 1110 individuals, so all genes with presences < 1110 are variable.

```{r}
res_tibb <- res_tibb %>% mutate(type = case_when(presences == 1110 ~ 'core',
                                     TRUE ~ 'variable'))
```

```{r}
nbs_joined <- left_join(class_dict[['NBS']], res_tibb, by = c('Name'='names'))
```

There are `r nrow(nbs_joined)` NLR genes out of which `r nbs_joined %>% filter(type == 'core') %>% nrow()` are core and `r nbs_joined %>% filter(type == 'variable') %>% nrow()` are variable.


```{r}
rlk_joined <- left_join(class_dict[['RLK']], res_tibb, by = c('Name'='names'))
```

There are `r nrow(rlk_joined)` RLK genes out of which `r rlk_joined %>% filter(type == 'core') %>% nrow()` are core and `r rlk_joined %>% filter(type == 'variable') %>% nrow()` are variable.


```{r}
rlp_joined <- left_join(class_dict[['RLP']], res_tibb, by = c('Name'='names'))
```

There are `r nrow(rlp_joined)` RLK genes out of which `r rlp_joined %>% filter(type == 'core') %>% nrow()` are core and `r  rlp_joined %>% filter(type == 'variable') %>% nrow()` are variable.

