---
title: "total_numbers"
author: "Philipp Bayer"
date: "2020-09-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

This is the same analysis as first-analysis, but with total numbers, not percentages genes lost


```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse)
library(patchwork)
library(ggsci)
library(dabestr)
library(dabestr)
library(cowplot)
library(ggsignif)
library(ggforce)

theme_set(theme_cowplot())
```

## Introduction


```{r}
npg_col = pal_npg("nrc")(9)
col_list <- c(`Wild-type`=npg_col[8],
   Landrace = npg_col[3],
  `Old cultivar`=npg_col[2],
  `Modern cultivar`=npg_col[4])

pav_table <- read_tsv('./data/soybean_pan_pav.matrix_gene.txt.gz')
```

## NBS part

Let's pull the NBS genes from the table
```{r}
nbs <- read_tsv('./data/Lee.NBS.candidates.lst', col_names = c('Name', 'Class'))
nbs
# have to remove the .t1s 
nbs$Name <- gsub('.t1','', nbs$Name)
```


```{r}
nbs_pav_table <- pav_table %>% filter(Individual %in% nbs$Name)
```
### Modern vs Old gene loss

```{r}
groups <- read_csv('./data/Table_of_cultivar_groups.csv')
groups
```

Which genes are present more or less in old / modern cultivars?

```{r big_norm_count}
big_norm_count <- tibble(
  name = character(),
  landrace = numeric(),
  Modern_cultivar = numeric(),
  Old_cultivar = numeric(),
  `Wild-type` = numeric()
)

groups_list <- split(groups$`Group in violin table`, groups$`Data-storage-ID`)

for( i in 1:nrow(nbs_pav_table) ) {
  this_gene <- nbs_pav_table[i,]
  groups_count <- list()
  total_groups_count <- list()
  for (x in seq_along(nbs_pav_table)){
    if ( x == 1) next
    thisind <- colnames(nbs_pav_table)[x]
    thisind_group <- groups_list[[thisind]]
    if( is.null(thisind_group) ) next # no group for this individual
    pavs <- this_gene[[x]] # either 1 or 0
    
    if ( thisind_group %in% names(groups_count)) {
      # count the number of present genes
      groups_count[[thisind_group]] <- groups_count[[thisind_group]] + pavs
      # count the total number of individuals for this group
      total_groups_count[[thisind_group]] <- total_groups_count[[thisind_group]] + 1
    } else {
      groups_count[[thisind_group]] <-  pavs
      total_groups_count[[thisind_group]] <- 1
    }
  }
  
  norm_group_count <- list()
  for (m in seq_along(groups_count)) {
    thisname <- names(groups_count)[m]
    norm_group_count[[thisname]] <- groups_count[[thisname]] / total_groups_count[[thisname]] * 100
  }
  norm_group_count$Individual <- this_gene$Individual
  big_norm_count <- rbind(big_norm_count, as_tibble(norm_group_count))

}

# wow, I DO write R like Python
```

Let's pull out the genes that are variable in any group

```{r}

var_norm_count <- big_norm_count %>% 
  filter(landrace != 100 & 
           Modern_cultivar != 100 & 
           Old_cultivar != 100 & 
           `Wild-type` != 100)

var_norm_count <- left_join(var_norm_count, nbs, by=c('Individual'='Name'))
var_norm_count$Mod_minus_Old <- var_norm_count$Modern_cultivar - var_norm_count$Old_cultivar
```

The top 20 genes reduced the most in modern cultivars compared with old cultivars:

```{r}
var_norm_count %>% 
  arrange(Mod_minus_Old) %>% 
  head(20) %>% 
  select(Individual, `Wild-type`, landrace, Old_cultivar, Modern_cultivar, Mod_minus_Old, Class) %>% 
  knitr::kable()
```

So these are the NLR genes selected against during soybean breeding.

Let's look at those without the TX ones:
```{r}
var_norm_count %>% 
  arrange(Mod_minus_Old) %>% 
  select(Individual, `Wild-type`, landrace, Old_cultivar, Modern_cultivar, Mod_minus_Old, Class) %>% 
  filter(Class != 'TX') %>% 
  head(20) %>% 
  knitr::kable()
```


The top 20 genes increased the most in modern cultivars:
```{r}
var_norm_count %>% 
  arrange(desc(Mod_minus_Old)) %>% 
  head(20) %>% 
  select(Individual, `Wild-type`, landrace, Old_cultivar, Modern_cultivar, Mod_minus_Old, Class) %>% 
  knitr::kable()
```

As these genes have relatively high percentages in WT they must have been re-introduced by using WT in the breeding process.


### Presence plotting per individual

```{r}
names <- c()
presences <- c()

for (i in seq_along(nbs_pav_table)){
  if ( i == 1) next
  thisind <- colnames(nbs_pav_table)[i]
  pavs <- nbs_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
nbs_res_tibb <- new_tibble(list(names = names, presences = presences))
```

OK what do these presence percentages look like?
```{r}
ggplot(data=nbs_res_tibb, aes(x=presences)) + geom_histogram(bins=25) 
```

On average, `r mean(nbs_res_tibb$presences)` of NBS genes are present in each individual.

Now let's join the table of presences to the four different types so we can group these numbers.

```{r}
nbs_joined_groups <- left_join(nbs_res_tibb, groups, by = c('names'='Data-storage-ID'))
```

```{r}
nbs_joined_groups$`Group in violin table` <- gsub('landrace', 'Landrace', nbs_joined_groups$`Group in violin table`)
nbs_joined_groups$`Group in violin table` <- gsub('Modern_cultivar', 'Modern cultivar', nbs_joined_groups$`Group in violin table`)
nbs_joined_groups$`Group in violin table` <- gsub('Old_cultivar', 'Old cultivar', nbs_joined_groups$`Group in violin table`)

nbs_joined_groups$`Group in violin table` <- factor(nbs_joined_groups$`Group in violin table`, levels=c(NA, 'Wild-type', 'Landrace', 'Old cultivar', 'Modern cultivar'))
```

```{r}
nbs_vio <- nbs_joined_groups %>% filter(!is.na(`Group in violin table`)) %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  geom_sina(alpha=0.5) +
  geom_smooth(aes(group=1), method='glm') +
  scale_fill_manual(values=col_list) +
  guides(fill = FALSE)

nbs_vio
```
```{r}
nbs_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_smooth(aes(group=1), method='lm', se = FALSE) +
  geom_jitter() +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE)
  

```

```{r}
nbs_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  group_by(`Group in violin table`) %>% 
  summarise(min_present = min(presences),
            max_present = max(presences),
            mean_present = mean(presences),
            median_present = median(presences),
            std_present = sd(presences)) %>% 
  knitr::kable()
```
## RLK part

Let's do the  same plot with RLKs


```{r}
rlk <- read_tsv('./data/Lee.RLK.candidates.lst', col_names = c('Name', 'Class', 'Subtype'))
rlk
# have to remove the .t1s 
rlk$Name <- gsub('.t1','', rlk$Name)
```

```{r}
rlk_pav_table <- pav_table %>% filter(Individual %in% rlk$Name)
rlk_pav_table
```

```{r}
names <- c()
presences <- c()

for (i in seq_along(rlk_pav_table)){
  if ( i == 1) next
  thisind <- colnames(rlk_pav_table)[i]
  pavs <- rlk_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
rlk_res_tibb <- new_tibble(list(names = names, presences = presences))
rlk_res_tibb
```

OK what do these presence percentages look like?
```{r}
ggplot(data=rlk_res_tibb, aes(x=presences)) + geom_histogram(bins=25) 
```

On average, `r mean(rlk_res_tibb$presences)`% of NBS genes are present in each individual.

Now let's join the table of presences to the four different types so we can group these numbers.

```{r}
rlk_joined_groups <- left_join(rlk_res_tibb, groups, by = c('names'='Data-storage-ID'))
```

```{r}
rlk_joined_groups$`Group in violin table` <- gsub('landrace', 'Landrace', rlk_joined_groups$`Group in violin table`)
rlk_joined_groups$`Group in violin table` <- gsub('Modern_cultivar', 'Modern cultivar', rlk_joined_groups$`Group in violin table`)
rlk_joined_groups$`Group in violin table` <- gsub('Old_cultivar', 'Old cultivar', rlk_joined_groups$`Group in violin table`)

rlk_joined_groups$`Group in violin table` <- factor(rlk_joined_groups$`Group in violin table`, levels=c(NA, 'Wild-type', 'Landrace', 'Old cultivar', 'Modern cultivar'))
```

```{r}
rlk_vio <- rlk_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  geom_sina(alpha=0.5) +
  geom_smooth(aes(group=1), method='lm', se = FALSE) +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE)
rlk_vio
```

```{r}
rlk_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  group_by(`Group in violin table`) %>% 
  summarise(min_present = min(presences),
            max_present = max(presences),
            mean_present = mean(presences),
            median_present = median(presences),
            std_present = sd(presences)) %>% 
  knitr::kable()
```
## RLP part

And now with RLPs



```{r}
rlp <- read_tsv('./data/Lee.RLP.candidates.lst', col_names = c('Name', 'Class', 'Subtype'))
# have to remove the .t1s 
rlp$Name <- gsub('.t1','', rlp$Name)
```

```{r}
rlp_pav_table <- pav_table %>% filter(Individual %in% rlp$Name)
```

```{r}
names <- c()
presences <- c()

for (i in seq_along(rlp_pav_table)){
  if ( i == 1) next
  thisind <- colnames(rlp_pav_table)[i]
  pavs <- rlp_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
rlp_res_tibb <- new_tibble(list(names = names, presences = presences))

```

OK what do these presence percentages look like?
```{r}
ggplot(data=rlp_res_tibb, aes(x=presences)) + geom_histogram(bins=25) 
```

On average, `r mean(rlp_res_tibb$presences)`% of NBS genes are present in each individual.

Now let's join the table of presences to the four different types so we can group these numbers.

```{r}
rlp_joined_groups <- left_join(rlp_res_tibb, groups, by = c('names'='Data-storage-ID'))
```

```{r}
rlp_joined_groups$`Group in violin table` <- gsub('landrace', 'Landrace', rlp_joined_groups$`Group in violin table`)
rlp_joined_groups$`Group in violin table` <- gsub('Modern_cultivar', 'Modern cultivar', rlp_joined_groups$`Group in violin table`)
rlp_joined_groups$`Group in violin table` <- gsub('Old_cultivar', 'Old cultivar', rlp_joined_groups$`Group in violin table`)

rlp_joined_groups$`Group in violin table` <- factor(rlp_joined_groups$`Group in violin table`, levels=c(NA, 'Wild-type', 'Landrace', 'Old cultivar', 'Modern cultivar'))
```

```{r}
rlp_vio <- rlp_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  geom_sina(alpha=0.5) +
    geom_smooth(aes(group=1), method='lm', se = FALSE) +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE)
rlp_vio
```

```{r}
rlp_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_jitter() +
  #geom_sina(alpha=0.5) +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE) +
  ylim(c(87, 100))
```

```{r}
rlp_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  group_by(`Group in violin table`) %>% 
  summarise(min_present = min(presences),
            max_present = max(presences),
            mean_present = mean(presences),
            median_present = median(presences),
            std_present = sd(presences)) %>% 
  knitr::kable()
```
## Plotting together

```{r}
nbs_vio + rlk_vio + rlp_vio
```

## Stats - Dabayes

I want to know whether the groups are statistically significantly different. First let's use dabestr


### NBS

Let's run dabestr first:

```{r}
nbs_multi.two.group.unpaired <- 
  nbs_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  dabest(`Group in violin table`, presences, 
         idx = list(c("Wild-type", "Landrace"),
                    c('Old cultivar', 'Modern cultivar')),
         paired = FALSE)
nbs_multi.two.group.unpaired
```

```{r}
nbs_multi.two.group.unpaired.meandiff <- mean_diff(nbs_multi.two.group.unpaired)
nbs_multi.two.group.unpaired.meandiff
```


```{r}
plot(nbs_multi.two.group.unpaired.meandiff, color.column=`Group in violin table`,
     rawplot.ylabel = 'Presence (%)', show.legend=FALSE)
```

### RLK

```{r}
rlk_multi.two.group.unpaired <- 
  rlk_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  dabest(`Group in violin table`, presences, 
         idx = list(c("Wild-type", "Landrace"),
                    c('Old cultivar', 'Modern cultivar')),
         paired = FALSE)
rlk_multi.two.group.unpaired
```

```{r}
rlk_multi.two.group.unpaired.meandiff <- mean_diff(rlk_multi.two.group.unpaired)
rlk_multi.two.group.unpaired.meandiff
```


```{r}
plot(rlk_multi.two.group.unpaired.meandiff, color.column=`Group in violin table`,
     rawplot.ylabel = 'Presence (%)', show.legend=FALSE)
```

No difference between old and modern cultivars!


### RLP

```{r}
rlp_multi.two.group.unpaired <- 
  rlp_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  dabest(`Group in violin table`, presences, 
         idx = list(c("Wild-type", "Landrace"),
                    c('Old cultivar', 'Modern cultivar')),
         paired = FALSE)
rlp_multi.two.group.unpaired
```

```{r}
rlp_multi.two.group.unpaired.meandiff <- mean_diff(rlp_multi.two.group.unpaired)
rlp_multi.two.group.unpaired.meandiff
```


```{r}
plot(rlp_multi.two.group.unpaired.meandiff, color.column=`Group in violin table`,
     rawplot.ylabel = 'Presence (%)', show.legend=FALSE)
```

Again, no difference between old and modern cultivars!



## Stats - classic t-test

### NBS

```{r}
nbs_joined_groups %>% 
  filter( !is.na(`PI-ID`) ) %>%
    ggplot(aes(x=`Group in violin table`, y = presences,
               fill = `Group in violin table`)) + 
  geom_boxplot() +
  scale_fill_manual(values = col_list) + 
  theme_minimal_hgrid() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12)) +
  geom_signif(comparisons = list(c('Wild-type', 'Landrace'),
                                 c('Old cultivar', 'Modern cultivar')), 
              map_signif_level = T) +
  guides(fill=FALSE) +
  ylab('Number of NLR genes') +
  xlab('Accession group')
```

### RLP
```{r}
rlp_joined_groups %>% 
  filter( !is.na(`PI-ID`) ) %>%
    ggplot(aes(x=`Group in violin table`, y = presences,
               fill = `Group in violin table`)) + 
  geom_boxplot() +
  scale_fill_manual(values = col_list) + 
  theme_minimal_hgrid() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12)) +
  geom_signif(comparisons = list(c('Wild-type', 'Landrace'),
                                 c('Old cultivar', 'Modern cultivar')), 
              map_signif_level = T) +
  guides(fill=FALSE) +
  ylab('Number of RLP genes') +
  xlab('Accession group')
```

### RLK

```{r}
rlk_joined_groups %>% 
  filter( !is.na(`PI-ID`) ) %>%
    ggplot(aes(x=`Group in violin table`, y = presences,
               fill = `Group in violin table`)) + 
  geom_boxplot() +
  scale_fill_manual(values = col_list) + 
  theme_minimal_hgrid() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12)) +
  geom_signif(comparisons = list(c('Wild-type', 'Landrace'),
                                 c('Old cultivar', 'Modern cultivar')), 
              map_signif_level = T) +
  guides(fill=FALSE) +
  ylab('Number of RLK genes') +
  xlab('Accession group')
```



## Linking with yield

Can we link the trajectory of NLR genes with the trajectory of yield across the history of soybean breeding? let's make a simple regression for now

```{r protein_join}
protein <- read_tsv('./data/protein_phenotype.txt')
protein_join <- left_join(nbs_res_tibb, protein, by=c('names'='Line')) %>% filter(!is.na(Protein))
```

```{r}
protein_join %>% ggplot(aes(x=presences, y=Protein)) + geom_hex() + geom_smooth() +
  xlab('NLR gene count')
```


```{r seed_join}
seed_weight <- read_tsv('./data/Seed_weight_Phenotype.txt', col_names = c('names', 'wt'))
seed_join <- left_join(nbs_res_tibb, seed_weight) %>% filter(!is.na(wt))
```

```{r}
seed_join %>% filter(wt > 5) %>%  ggplot(aes(x=presences, y=wt)) + geom_hex() + geom_smooth() +
  ylab('Seed weight') +
  xlab('NLR gene count')
```


```{r oil_join}
oil <- read_tsv('./data/oil_phenotype.txt')
oil_join <- left_join(nbs_res_tibb, oil, by=c('names'='Line')) %>% filter(!is.na(Oil))
```


```{r}
oil_join %>%  ggplot(aes(x=presences, y=Oil)) + geom_hex() + geom_smooth() +
  xlab('NLR gene count')
```


OK there are many, many outliers here. Clearly I'll have to do something fancier - for example, using the first two PCs as covariates might get rid of some of those outliers. 
