---
title: "total_numbers"
author: "Philipp Bayer"
date: "2020-09-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

This is the same analysis as first-analysis, but with total numbers, not percentages genes lost


```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse)
library(patchwork)
library(ggsci)
library(dabestr)
library(dabestr)
library(cowplot)
library(ggsignif)
library(ggforce)

theme_set(theme_cowplot())
```

## Introduction


```{r}
npg_col = pal_npg("nrc")(9)
col_list <- c(`Wild-type`=npg_col[8],
   Landrace = npg_col[3],
  `Old cultivar`=npg_col[2],
  `Modern cultivar`=npg_col[4])

pav_table <- read_tsv('./data/soybean_pan_pav.matrix_gene.txt.gz')
```

## NBS part

Let's pull the NBS genes from the table
```{r}
nbs <- read_tsv('./data/Lee.NBS.candidates.lst', col_names = c('Name', 'Class'))
nbs
# have to remove the .t1s 
nbs$Name <- gsub('.t1','', nbs$Name)
```


```{r}
nbs_pav_table <- pav_table %>% filter(Individual %in% nbs$Name)
```

```{r}
names <- c()
presences <- c()

for (i in seq_along(nbs_pav_table)){
  if ( i == 1) next
  thisind <- colnames(nbs_pav_table)[i]
  pavs <- nbs_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
nbs_res_tibb <- new_tibble(list(names = names, presences = presences))
```

OK what do these presence percentages look like?
```{r}
ggplot(data=nbs_res_tibb, aes(x=presences)) + geom_histogram(bins=25) 
```

On average, `r mean(nbs_res_tibb$presences)` of NBS genes are present in each individual.

Now let's join the table of presences to the four different types so we can group these numbers.

```{r}
nbs_groups <- read_csv('./data/Table_of_cultivar_groups.csv')
nbs_joined_groups <- left_join(nbs_res_tibb, nbs_groups, by = c('names'='Data-storage-ID'))
```

```{r}
nbs_joined_groups$`Group in violin table` <- gsub('landrace', 'Landrace', nbs_joined_groups$`Group in violin table`)
nbs_joined_groups$`Group in violin table` <- gsub('Modern_cultivar', 'Modern cultivar', nbs_joined_groups$`Group in violin table`)
nbs_joined_groups$`Group in violin table` <- gsub('Old_cultivar', 'Old cultivar', nbs_joined_groups$`Group in violin table`)

nbs_joined_groups$`Group in violin table` <- factor(nbs_joined_groups$`Group in violin table`, levels=c(NA, 'Wild-type', 'Landrace', 'Old cultivar', 'Modern cultivar'))
```

```{r}
nbs_vio <- nbs_joined_groups %>% filter(!is.na(`Group in violin table`)) %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  geom_sina(alpha=0.5) +
  geom_smooth(aes(group=1), method='glm') +
  scale_fill_manual(values=col_list) +
  guides(fill = FALSE)

nbs_vio
```
```{r}
nbs_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_smooth(aes(group=1), method='lm', se = FALSE) +
  geom_jitter() +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE)
  

```

```{r}
nbs_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  group_by(`Group in violin table`) %>% 
  summarise(min_present = min(presences),
            max_present = max(presences),
            mean_present = mean(presences),
            median_present = median(presences),
            std_present = sd(presences)) %>% 
  knitr::kable()
```
## RLK part

Let's do the  same plot with RLKs


```{r}
rlk <- read_tsv('./data/Lee.RLK.candidates.lst', col_names = c('Name', 'Class', 'Subtype'))
rlk
# have to remove the .t1s 
rlk$Name <- gsub('.t1','', rlk$Name)
```

```{r}
rlk_pav_table <- pav_table %>% filter(Individual %in% rlk$Name)
rlk_pav_table
```

```{r}
names <- c()
presences <- c()

for (i in seq_along(rlk_pav_table)){
  if ( i == 1) next
  thisind <- colnames(rlk_pav_table)[i]
  pavs <- rlk_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
rlk_res_tibb <- new_tibble(list(names = names, presences = presences))
rlk_res_tibb
```

OK what do these presence percentages look like?
```{r}
ggplot(data=rlk_res_tibb, aes(x=presences)) + geom_histogram(bins=25) 
```

On average, `r mean(rlk_res_tibb$presences)`% of NBS genes are present in each individual.

Now let's join the table of presences to the four different types so we can group these numbers.

```{r}
groups <- read_csv('./data/Table_of_cultivar_groups.csv')
groups
rlk_joined_groups <- left_join(rlk_res_tibb, groups, by = c('names'='Data-storage-ID'))
```

```{r}
rlk_joined_groups$`Group in violin table` <- gsub('landrace', 'Landrace', rlk_joined_groups$`Group in violin table`)
rlk_joined_groups$`Group in violin table` <- gsub('Modern_cultivar', 'Modern cultivar', rlk_joined_groups$`Group in violin table`)
rlk_joined_groups$`Group in violin table` <- gsub('Old_cultivar', 'Old cultivar', rlk_joined_groups$`Group in violin table`)

rlk_joined_groups$`Group in violin table` <- factor(rlk_joined_groups$`Group in violin table`, levels=c(NA, 'Wild-type', 'Landrace', 'Old cultivar', 'Modern cultivar'))
```

```{r}
rlk_vio <- rlk_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  geom_sina(alpha=0.5) +
  geom_smooth(aes(group=1), method='lm', se = FALSE) +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE)
rlk_vio
```

```{r}
rlk_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  group_by(`Group in violin table`) %>% 
  summarise(min_present = min(presences),
            max_present = max(presences),
            mean_present = mean(presences),
            median_present = median(presences),
            std_present = sd(presences)) %>% 
  knitr::kable()
```
## RLP part

And now with RLPs



```{r}
rlp <- read_tsv('./data/Lee.RLP.candidates.lst', col_names = c('Name', 'Class', 'Subtype'))
# have to remove the .t1s 
rlp$Name <- gsub('.t1','', rlp$Name)
```

```{r}
rlp_pav_table <- pav_table %>% filter(Individual %in% rlp$Name)
```

```{r}
names <- c()
presences <- c()

for (i in seq_along(rlp_pav_table)){
  if ( i == 1) next
  thisind <- colnames(rlp_pav_table)[i]
  pavs <- rlp_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
rlp_res_tibb <- new_tibble(list(names = names, presences = presences))

```

OK what do these presence percentages look like?
```{r}
ggplot(data=rlp_res_tibb, aes(x=presences)) + geom_histogram(bins=25) 
```

On average, `r mean(rlp_res_tibb$presences)`% of NBS genes are present in each individual.

Now let's join the table of presences to the four different types so we can group these numbers.

```{r}
groups <- read_csv('./data/Table_of_cultivar_groups.csv')

rlp_joined_groups <- left_join(rlp_res_tibb, groups, by = c('names'='Data-storage-ID'))
```

```{r}
rlp_joined_groups$`Group in violin table` <- gsub('landrace', 'Landrace', rlp_joined_groups$`Group in violin table`)
rlp_joined_groups$`Group in violin table` <- gsub('Modern_cultivar', 'Modern cultivar', rlp_joined_groups$`Group in violin table`)
rlp_joined_groups$`Group in violin table` <- gsub('Old_cultivar', 'Old cultivar', rlp_joined_groups$`Group in violin table`)

rlp_joined_groups$`Group in violin table` <- factor(rlp_joined_groups$`Group in violin table`, levels=c(NA, 'Wild-type', 'Landrace', 'Old cultivar', 'Modern cultivar'))
```

```{r}
rlp_vio <- rlp_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_violin(draw_quantiles = c(0.5)) +
  geom_sina(alpha=0.5) +
    geom_smooth(aes(group=1), method='lm', se = FALSE) +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE)
rlp_vio
```

```{r}
rlp_joined_groups %>% filter(`Group in violin table` != 'NA') %>% 
  ggplot(aes(y=presences, x=`Group in violin table`, fill=`Group in violin table`)) + 
  geom_jitter() +
  #geom_sina(alpha=0.5) +
  scale_fill_manual(values=col_list)+
  guides(fill = FALSE) +
  ylim(c(87, 100))
```

```{r}
rlp_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  group_by(`Group in violin table`) %>% 
  summarise(min_present = min(presences),
            max_present = max(presences),
            mean_present = mean(presences),
            median_present = median(presences),
            std_present = sd(presences)) %>% 
  knitr::kable()
```
## Plotting together

```{r}
nbs_vio + rlk_vio + rlp_vio
```

## Stats - Dabayes

I want to know whether the groups are statistically significantly different. First let's use dabestr


### NBS

Let's run dabestr first:

```{r}
nbs_multi.two.group.unpaired <- 
  nbs_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  dabest(`Group in violin table`, presences, 
         idx = list(c("Wild-type", "Landrace"),
                    c('Old cultivar', 'Modern cultivar')),
         paired = FALSE)
nbs_multi.two.group.unpaired
```

```{r}
nbs_multi.two.group.unpaired.meandiff <- mean_diff(nbs_multi.two.group.unpaired)
nbs_multi.two.group.unpaired.meandiff
```


```{r}
plot(nbs_multi.two.group.unpaired.meandiff, color.column=`Group in violin table`,
     rawplot.ylabel = 'Presence (%)', show.legend=FALSE)
```

### RLK

```{r}
rlk_multi.two.group.unpaired <- 
  rlk_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  dabest(`Group in violin table`, presences, 
         idx = list(c("Wild-type", "Landrace"),
                    c('Old cultivar', 'Modern cultivar')),
         paired = FALSE)
rlk_multi.two.group.unpaired
```

```{r}
rlk_multi.two.group.unpaired.meandiff <- mean_diff(rlk_multi.two.group.unpaired)
rlk_multi.two.group.unpaired.meandiff
```


```{r}
plot(rlk_multi.two.group.unpaired.meandiff, color.column=`Group in violin table`,
     rawplot.ylabel = 'Presence (%)', show.legend=FALSE)
```

No difference between old and modern cultivars!


### RLP

```{r}
rlp_multi.two.group.unpaired <- 
  rlp_joined_groups %>% filter(!is.na(`PI-ID`)) %>% 
  dabest(`Group in violin table`, presences, 
         idx = list(c("Wild-type", "Landrace"),
                    c('Old cultivar', 'Modern cultivar')),
         paired = FALSE)
rlp_multi.two.group.unpaired
```

```{r}
rlp_multi.two.group.unpaired.meandiff <- mean_diff(rlp_multi.two.group.unpaired)
rlp_multi.two.group.unpaired.meandiff
```


```{r}
plot(rlp_multi.two.group.unpaired.meandiff, color.column=`Group in violin table`,
     rawplot.ylabel = 'Presence (%)', show.legend=FALSE)
```

Again, no difference between old and modern cultivars!



## Stats - classic t-test

### NBS

```{r}
nbs_joined_groups %>% 
  filter( !is.na(`PI-ID`) ) %>%
    ggplot(aes(x=`Group in violin table`, y = presences,
               fill = `Group in violin table`)) + 
  geom_boxplot() +
  scale_fill_manual(values = col_list) + 
  theme_minimal_hgrid() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12)) +
  geom_signif(comparisons = list(c('Wild-type', 'Landrace'),
                                 c('Old cultivar', 'Modern cultivar')), 
              map_signif_level = T) +
  guides(fill=FALSE) +
  ylab('Number of NLR genes') +
  xlab('Accession group')
```

### RLP
```{r}
rlp_joined_groups %>% 
  filter( !is.na(`PI-ID`) ) %>%
    ggplot(aes(x=`Group in violin table`, y = presences,
               fill = `Group in violin table`)) + 
  geom_boxplot() +
  scale_fill_manual(values = col_list) + 
  theme_minimal_hgrid() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12)) +
  geom_signif(comparisons = list(c('Wild-type', 'Landrace'),
                                 c('Old cultivar', 'Modern cultivar')), 
              map_signif_level = T) +
  guides(fill=FALSE) +
  ylab('Number of RLP genes') +
  xlab('Accession group')
```

### RLK

```{r}
rlk_joined_groups %>% 
  filter( !is.na(`PI-ID`) ) %>%
    ggplot(aes(x=`Group in violin table`, y = presences,
               fill = `Group in violin table`)) + 
  geom_boxplot() +
  scale_fill_manual(values = col_list) + 
  theme_minimal_hgrid() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12)) +
  geom_signif(comparisons = list(c('Wild-type', 'Landrace'),
                                 c('Old cultivar', 'Modern cultivar')), 
              map_signif_level = T) +
  guides(fill=FALSE) +
  ylab('Number of RLK genes') +
  xlab('Accession group')
```
