---
title: "Linking yield with NLR PAV"
author: "Philipp Bayer"
date: "2020-09-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(message = FALSE) 
library(tidyverse)
library(patchwork)
library(sjPlot)
library(ggsci)
library(dabestr)
library(dabestr)
library(cowplot)
library(ggsignif)
library(ggforce)
library(lme4)
library(directlabels)
library(lmerTest)
library(dotwhisker)
library(pals)
theme_set(theme_cowplot())
library(RColorBrewer)
library(countrycode)
library(broom)
```


# Data loading

```{r}
npg_col = pal_npg("nrc")(9)
col_list <- c(`Wild`=npg_col[8],
   Landrace = npg_col[3],
  `Old cultivar`=npg_col[2],
  `Modern cultivar`=npg_col[4])

pav_table <- read_tsv('./data/soybean_pan_pav.matrix_gene.txt.gz')
```


```{r}
nbs <- read_tsv('./data/Lee.NBS.candidates.lst', col_names = c('Name', 'Class'))
nbs
# have to remove the .t1s 
nbs$Name <- gsub('.t1','', nbs$Name)
nbs_pav_table <- pav_table %>% filter(Individual %in% nbs$Name)
```


```{r}
names <- c()
presences <- c()

for (i in seq_along(nbs_pav_table)){
  if ( i == 1) next
  thisind <- colnames(nbs_pav_table)[i]
  pavs <- nbs_pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
nbs_res_tibb <- new_tibble(list(names = names, presences = presences))
```

```{r}
# let's make the same table for all genes too
names <- c()
presences <- c()

for (i in seq_along(pav_table)){
  if ( i == 1) next
  thisind <- colnames(pav_table)[i]
  pavs <- pav_table[[i]]
  presents <- sum(pavs)
  names <- c(names, thisind)
  presences <- c(presences, presents)
}
res_tibb <- new_tibble(list(names = names, presences = presences))
```


```{r}
groups <- read_csv('./data/Table_of_cultivar_groups.csv')
groups <- groups %>% dplyr::rename(Group = `Group in violin table`)
groups <- groups %>% 
  mutate(Group = str_replace_all(Group, 'landrace', 'Landrace')) %>%
  mutate(Group = str_replace_all(Group, 'Old_cultivar', 'Old cultivar')) %>%
  mutate(Group = str_replace_all(Group, 'Modern_cultivar', 'Modern cultivar')) %>%
  mutate(Group = str_replace_all(Group, 'Wild-type', 'Wild'))

groups$Group <-
  factor(
    groups$Group,
    levels = c('Wild',
               'Landrace',
               'Old cultivar',
               'Modern cultivar')
  )

nbs_joined_groups <-
  inner_join(nbs_res_tibb, groups, by = c('names' = 'Data-storage-ID'))
all_joined_groups <-
    inner_join(res_tibb, groups, by = c('names' = 'Data-storage-ID'))

```

```{r}
country <- read_csv('./data/Cultivar_vs_country.csv')


names(country) <- c('names', 'PI-ID', 'Country')

# fix weird ND issue
country <- country %>% mutate(Country = na_if(Country, 'ND'))

```

# Linking with yield

Can we link the trajectory of NLR genes with the trajectory of yield across the history of soybean breeding? let's make a simple regression for now

## Yield

```{r yield_join}

yield <- read_tsv('./data/yield.txt')
yield_join <- inner_join(nbs_res_tibb, yield, by=c('names'='Line'))
```

## Protein

```{r protein_join}
protein <- read_tsv('./data/protein_phenotype.txt')
protein_join <- left_join(nbs_res_tibb, protein, by=c('names'='Line')) %>% filter(!is.na(Protein))
```

## Seed weight

Let's look at seed weight:

```{r seed_join}
seed_weight <- read_tsv('./data/Seed_weight_Phenotype.txt', col_names = c('names', 'wt'))
seed_join <- left_join(nbs_res_tibb, seed_weight) %>% filter(!is.na(wt))
```

## Oil content
And now let's look at the oil phenotype:

```{r oil_join}
oil <- read_tsv('./data/oil_phenotype.txt')
oil_join <- left_join(nbs_res_tibb, oil, by=c('names'='Line')) %>% filter(!is.na(Oil))
oil_join
```


# Basic lm()

These results form the basis of the paper.

## Yield

```{r}
yield_nbs_joined_groups <- nbs_joined_groups %>% inner_join(yield_join, by = 'names') 
yield_nbs_joined_groups$Yield2 <-scale(yield_nbs_joined_groups$Yield, center=T, scale=T)

yield_all_joined_groups <- all_joined_groups %>% inner_join(yield_join, by = 'names')
```

```{r}
yield_country_nbs_joined_groups <- yield_nbs_joined_groups %>% inner_join(country)
yield_country_all_joined_groups <- yield_all_joined_groups %>% inner_join(country)
yield_country_nbs_joined_groups$Count <- yield_country_nbs_joined_groups$presences.y
yield_country_all_joined_groups$Count <- yield_country_all_joined_groups$presences.y

yield_country_all_joined_groups %>% count(Country)
```

```{r}
lm_model <- lm(Yield ~ Count +  Group, data = yield_country_nbs_joined_groups) 
summary(lm_model)
```

```{r}
clm_model <- lm(Yield ~ Count +  Group + Country, data = yield_country_nbs_joined_groups) 
summary(clm_model)
yieldmod <- clm_model
```

```{r}
interlm_model <- lm(Yield ~ Count * Group, data = yield_country_nbs_joined_groups) 
summary(interlm_model)
```


```{r}
tab_model(clm_model, p.val='kr', digits=3)
```

```{r}
newdat <-cbind(yield_country_all_joined_groups %>% filter(!is.na(Country)), pred = predict(clm_model))
```

```{r}
# I want only every second, stronger color of the Paired scheme
mycol <- brewer.pal(n = 12, name = "Paired")[seq(2, 12, 2)]
```

```{r}
newdat %>% mutate(
  Country2 = case_when (
  Country == 'USA' ~ 'USA',
  Country == 'China' ~ 'China',
  Country == 'Korea' ~ 'Korea',
  Country == 'Japan' ~ 'Japan',
  Country == 'Russia' ~ 'Russia',
  TRUE ~ 'Rest'
  )
  ) %>%
  mutate(Country2 = factor(
  Country2,
  levels = c('China', 'Japan', 'Korea', 'Russia', 'USA', 'Rest')
  )) %>%
  ggplot(aes(x = Count, y = pred, color = Country2)) +
  facet_wrap( ~ Group, nrow = 1) +  
  geom_point(aes(y = Yield, color = Country2),
    alpha = 0.8,
    size = 2) +
  geom_line(aes(y=pred, group=interaction(Group, Country),colour=Country2), size=1.5)+
  theme_minimal_hgrid() +
  xlab('Gene count') +
  ylab(expression(paste('Yield [Mg ', ha ^ -1, ']'))) +
  scale_color_manual(values = mycol) +
  #xlim(c(47.900, 49.700)) +
  labs(color = "Country") +
  theme(panel.spacing = unit(0.9, "lines"),
  axis.text.x = element_text(size = 10))
```

Oh wow, that looks very similar.

```{r}
plot_model(clm_model, type = "pred", terms = c("Count", "Group")) +
  theme_minimal_hgrid() +
  #scale_fill_manual(values = col_list) +
  #scale_color_manual(values = col_list) +
  xlab('NLR count') + 
  ylab((expression(paste('Yield [Mg ', ha^-1, ']')))) +
  theme(plot.title=element_blank())
```

```{r}
qqnorm(resid(clm_model))
qqline(resid(clm_model))
```

```{r}
plot(resid(clm_model))
```

```{r}
plot_model(clm_model, show.values = TRUE, value.offset = .3, terms=c('GroupOld cultivar', 'Count', 'GroupModern cultivar'))
```

```{r fig.height=8, fig.width=8}
plot_model(clm_model, type = "std", show.values = TRUE, value.offset = .3,)
```

```{r}
plot_model(clm_model, type = "std", show.values = TRUE, value.offset = .3,, terms=c('GroupOld cultivar', 'Count', 'GroupModern cultivar'))
```

## Protein
```{r}
protein_nbs_joined_groups <- nbs_joined_groups %>% inner_join(protein_join, by = 'names') 
#protein_nbs_joined_groups$Protein2 <- scale(protein_nbs_joined_groups$Protein, center=T, scale=T)
protein_country_nbs_joined_groups <- protein_nbs_joined_groups %>% inner_join(country)
#protein_country_nbs_joined_groups <- rename(protein_country_nbs_joined_groups, Group=`Group in violin table`)
protein_country_nbs_joined_groups <- protein_country_nbs_joined_groups %>% filter(Group != 'Wild')
protein_country_nbs_joined_groups <- protein_country_nbs_joined_groups %>% rename(Count = presences.x)
```


```{r}
lm_model <- lm(Protein ~ Count +  Group, data = protein_country_nbs_joined_groups) 
summary(lm_model)
```

```{r}
clm_model <- lm(Protein ~ Count +  Group + Country, data = protein_country_nbs_joined_groups) 
summary(clm_model)
protmod <- clm_model
```

## Seed weight

```{r}
seed_nbs_joined_groups <- nbs_joined_groups %>% inner_join(seed_join, by = 'names') 
#seed_nbs_joined_groups$wt2 <- scale(seed_nbs_joined_groups$wt, center=T, scale=T)
seed_country_nbs_joined_groups <- seed_nbs_joined_groups %>% inner_join(country)
#seed_country_nbs_joined_groups <- rename(seed_country_nbs_joined_groups, Group = `Group in violin table`)
seed_country_nbs_joined_groups <- seed_country_nbs_joined_groups %>% filter(Group != 'Wild')
seed_country_nbs_joined_groups <- seed_country_nbs_joined_groups %>% rename(Count = presences.x)
```

```{r}
lm_model <- lm(wt ~ Count +  Group, data = seed_country_nbs_joined_groups) 
summary(lm_model)
```

```{r}
clm_model <- lm(wt ~ Count +  Group + Country, data = seed_country_nbs_joined_groups) 
summary(clm_model)
seedmod <- clm_model
```

Nothing at all.

## Oil content

```{r}
oil_nbs_joined_groups <- nbs_joined_groups %>% inner_join(oil_join, by = 'names') 
oil_nbs_joined_groups$Oil2 <- scale(oil_nbs_joined_groups$Oil, center=T, scale=T)
oil_country_nbs_joined_groups <- oil_nbs_joined_groups %>% inner_join(country)
oil_country_nbs_joined_groups <- oil_country_nbs_joined_groups %>% filter(Group != 'Wild')
oil_country_nbs_joined_groups <- oil_country_nbs_joined_groups %>% rename(Count = presences.x)
```

```{r}
lm_model <- lm(Oil ~ Count +  Group, data = oil_country_nbs_joined_groups) 
summary(lm_model)
```

Interesting, the simple model has p< 0.05 for oil content!

```{r}
clm_model <- lm(Oil ~ Count +  Group + Country, data = oil_country_nbs_joined_groups) 
summary(clm_model)
oilmod <- clm_model
```

p = 0.07 for Count + Group + Country in the model


## All models together

```{r}
tab_model(yieldmod, oilmod, protmod, seedmod, digits=3 )
```

## The yield model for all genes

```{r}
yield_country_all_joined_groups <- yield_country_all_joined_groups %>% rename(`All genes` = presences.x)
```

```{r}
clm <- lm(Yield ~ `All genes` + Group + Country, data=yield_country_all_joined_groups)
summary(clm)
```
Good, p = 0.78

```{r}
tab_model(clm, p.val='kr', digits=3)
```

## Continents, not countries 
```{r}

yield_country_nbs_joined_groups$continent <- countrycode(sourcevar = yield_country_nbs_joined_groups$Country,
                                                         origin = 'country.name',
                                                         destination = 'continent')

yield_country_nbs_joined_groups <- yield_country_nbs_joined_groups %>% mutate(continent2 = case_when (
  Country == 'USA' ~ 'North America',
  Country == 'Canada' ~ 'North America',
  continent == 'Americas' ~ 'South America',
  TRUE ~ continent
  )) 
```

```{r}
clm <- lm(Yield ~ Count + Group + continent2, data=yield_country_nbs_joined_groups)
summary(clm)
```

```{r}
yield_country_nbs_joined_groups_nonacont <- yield_country_nbs_joined_groups %>% filter(!is.na(continent2))
newdat <- cbind(yield_country_nbs_joined_groups_nonacont, pred = predict(clm))
```


```{r}
newdat %>% 
  ggplot(aes(x = Count, y = pred, color = continent2)) +
  facet_wrap( ~ Group, nrow = 1) +  
  geom_point(aes(y = Yield, color = continent2),
    alpha = 0.8,
    size = 2) +
  geom_line(aes(y=pred, group=interaction(Group, continent2),colour=continent2), size=1.5)+
  theme_minimal_hgrid() +
  xlab('Gene count') +
  ylab(expression(paste('Yield [Mg ', ha ^ -1, ']'))) +
  scale_color_manual(values = mycol) +
  #xlim(c(47.900, 49.700)) +
  labs(color = "Continent") +
  theme(panel.spacing = unit(0.9, "lines"),
  axis.text.x = element_text(size = 10))
```

## Finally, a regression by continent

I remove Oceania as there are very little individuals in there.

```{r}
by_continent_model <- function(df) {
  lm(Yield ~ Count + Group, data = df)
}
```

```{r}
with_models <- yield_country_nbs_joined_groups_nonacont %>% 
  filter(!continent2 %in% c('Africa', 'South America', 'Oceania')) %>% 
  group_by(continent2) %>% 
  nest() %>%
  mutate(models = map(data, by_continent_model))
```


```{r}
with_models %>% mutate(glance = map(models, broom::glance)) %>% 
  unnest(glance)
```

Let's pull out the data

```{r}
with_models %>% filter(continent2 == 'Asia') %>% unnest(data)
```

```{r}
plots <- with_models %>% 
  mutate(plots = map2(data, continent2, 
                      ~ ggplot(data =.x, aes(x=presences.x, y=Yield)) + 
                        geom_point() +
                        geom_smooth(method = "lm")
                      )
         )
```

I have no idea how to use standard facet_wrap with this.

```{r}
print(plots$plots)
```
```{r}
plots %>% unnest(data) %>% group_by(continent2) %>% 
  do(augment(.$models[[1]]))
```

